---
title: "Anomaly Detection IP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Defining the Question?
## Specifying the Question
You are a Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax)

Perform Anomaly detection on the sales dataset

## Defining the metrics of success
Perform data cleaning, Exploratory data analysis and Anomaly detection effectively in R

## Understanding the Context
Anomaly detection is collection of techniques designed to identify unusual data points. It helps identify unexpected data points in datasets.There are many applications of anomaly detections such as fraud and intrusion detection, system health monitoring and ecosystem disturbance monitoring. In our case we will be using anomaly detection to detect fraud.

## Recording the Experimental Design
 - Reading the Data
 - Checking the Data
 - Tidying the Data
 - Exploratory Data Analysis
 - Perform Anomaly Detection

## Data Relevance
The dataset provided was relevant for the objective we set for our project.



# Reading the Dataset
```{r}
# Loading the Dataset
sf = read.csv('Supermarket_Sales_Forecasting - Sales.csv')

```



# Checking the Dataset
```{r}
# Checking the shape of the dataset
dim(sf)

```
The dataset contains 1000 rows and 2 columns.


```{r}
# Checking the class of our dataset
class(sf)

```



```{r}
# Previewing the top of the Dataset
head(sf)

```


```{r}
# Previewing the bottom of the Dataset
tail(sf)

```


```{r}
# Checking the Datatypes of the columns
str(sf)

```
The dataset contains 1 column of datatype num and 1 column of datatype character


```{r}
# Checking for Duplicates
anyDuplicated(sf)

```
There were no duplicates in our dataset.



# Tidying the Dataset
## Missing values
```{r}
# Checking for missing values
colSums(is.na(sf))

```


## Outliers
```{r}
# Checking for Outliers in the sales column
boxplot(sf$Sales)

```


```{r}
# Checking the list of Outliers in the Sales column
boxplot.stats(sf$Sales)$out

```
There are 9 Outliers in the Sales column.
The Outliers in the column was removed because they had little effect on the data.



# Exploratory Data Analysis
### Mean
```{r}
mean(sf$Sales)

```


### Median
```{r}
median(sf$Sales)

```


### Maximum
```{r}
max(sf$Sales)

```


### Minimum
```{r}
min(sf$Sales)

```


### Range
```{r}
range(sf$Sales)

```


### Quantile
```{r}
quantile(sf$Sales)

```


### Interquantile Range
```{r}
IQR(sf$Sales)

```


### Variance
```{r}
var(sf$Sales)

```


### Standard Deviation
```{r}
sd(sf$Sales)

```


### Skewness
```{r}
library(e1071)
skewness(sf$Sales)

```


### Kurtosis
```{r}
kurtosis(sf$Sales)

```


### Histogram
```{r}
# Plotting the histogram of sales
hist(sf$Sales)

```


```{r}
# Plotting barplot of Dates 
barplot(table(sf$Date),
main = "Bargraph of Dates",
xlab = "Weekend",
ylab = "Freq",
col = topo.colors(5),
horiz = FALSE)

```

### Observations
 - Most sales ranged from 100 - 400
 - Maximum sale was 1042.65
 - Minimum sale was 10.6785
 - The mean sales were 322.9667

# Implementing the Solution
## Anomaly Detection
```{r}
# Load tidyverse and anomalize libraries
# 
library(tidyverse)
library(anomalize)

```


```{r}
# Checking the datatype of Date column
class(sf$Date)

```


```{r}
# Changing the Date column from character to date format
library(lubridate)
sf$Date <- as.Date(sf$Date, format="%m/%d/%Y")

```


```{r}
# Ordering the dataframe according to the dates
sf1 <- sf[order(sf$Date),]

```

```{r}
# Previewing the new dataframe
head(sf1)

```


```{r}
# Checking the datatype of Date column
class(sf1$Date)

```


```{r}
# Changing the dataset from dataframe format to tibble
sf_tbl = as_tibble(sf1)
head(sf_tbl)

```


```{r}
# Detecting the anomalies and plotting them.
sf_tbl %>%
  group_by(Date) %>%
  summarise(Sales = sum(Sales)) %>%
  time_decompose(Sales) %>%
  anomalize(remainder) %>%
  time_recompose() %>%
  plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)

```
There were no anomalies detected from the sales throughout the period the sales were recorded. Since there were no anomalies fraud did not take place in the sales.