---
title: "Association Analysis IP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Defining the Question?
## Specifying the Question
You are a Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax)

Perform Associative analysis.

## Defining the metrics of success
Perform data cleaning, Exploratory data analysis and Associative analysis effectively in R

## Understanding the Context
Companies require association rules to identify relationships between items in the in sales. Associative analysis is used to find associations between the items that were purchased and as a result derive association rules that identify the items and co-occurrences of different items that appear with the greatest frequencies.

## Recording the Experimental Design
 - Reading the Data
 - Checking the Data
 - Tidying the Data
 - Exploratory Data Analysis
 - Performing Associative analysis

## Data Relevance
The dataset provided was relevant for the objective we set for our project.



# Reading the Dataset
```{r}
# Loading the Dataset
sales = read.csv('Supermarket_Sales_Dataset II.csv')
```



# Checking the Dataset
```{r}
# Checking the shape of the dataset
dim(sales)
```
The dataset contains 7500 rows and 20 columns.


```{r}
# Previewing the top of the Dataset
head(sales)
```


```{r}
# Previewing the bottom of the Dataset
tail(sales)
```


```{r}
# Checking the Datatypes of the columns
str(sales)
```
The dataset contains 19 columns of datatype character and 1 column of datatype logical.


```{r}
# Checking for Duplicates
anyDuplicated(sales)
```
There were 34 duplicates in our dataset.


```{r}
# Dropping the duplicates in the dataset
sales = unique(sales)
```


# Tidying the Dataset
## Missing values
```{r}
# Checking for missing values
colSums(is.na(sales))
```

```{r}
# Omitting the column of the null values
sales = sales[, -20]
head(sales)
```



# Implementing the Solution
## Association Rules
```{r}
# Loading the arules library
library(arules)
```


```{r}
# Converting our dataset to class transaction.
sls<-read.transactions('Supermarket_Sales_Dataset II.csv')
sls
```


```{r}
# Verifying the class of our dataframe
class(sls)
```


```{r}
# Previewing the first 5 transactions of our datasets
#
inspect(sls[1:5])
```


```{r}
# Another method of previewing our dataset
# 
items<-as.data.frame(itemLabels(sls))
colnames(items) <- "Item"
head(items, 10)    
```


```{r}
# Getting the summary of our transactions in the dataset.
# 
summary(sls)
```


```{r}
# Exploring the frequency of some articles 
# 
itemFrequency(sls[, 8:10],type = "absolute")
round(itemFrequency(sls[, 8:10],type = "relative")*100,2)
```


```{r}
# 
par(mfrow = c(1, 2))

# plotting the frequency of items
itemFrequencyPlot(sls, topN = 10,col="darkgreen")

# Plotting items whose relative importance is at least 10% 
itemFrequencyPlot(sls, support = 0.1,col="darkred")
```


```{r}
# Building a apriori model with Min Support as 0.001 and confidence as 0.8
# 
rls <- apriori (sls, parameter = list(supp = 0.001, conf = 0.8))

rls

```


```{r}
# Getting the summary of the rules created
# 
summary(rls)
```


```{r}
# Observing rules built in our model
# 
inspect(rls[1:5])

```


```{r}
# Ordering these rules by a lift criteria
# 
rls<-sort(rls, by="lift", decreasing=TRUE)
inspect(rls[1:5])

```


```{r}
# Ordering these rules by a support criteria
rls<-sort(rls, by="support", decreasing=TRUE)
inspect(rls[1:5])
```


```{r}
# Ordering these rules by a confidence criteria
rls<-sort(rls, by="confidence", decreasing=TRUE)
inspect(rls[1:5])
```


## Observations
 - The most frequent items are tea, wheat, mineral, fat, and yogurt.
 - The model gives 271 set of rules.
 - The rules have to 4 items per subsets.
 - For every cookies and low someone buys there is a 100% chance they will buy yogurt and fat.


# Challenging the Solution
The solution can be challenged by building an apriori model with different support and confidence. This tends to reduce the number of rules.
```{r}
# Building a apriori model with Min Support as 0.002 and confidence as 0.6.
#
rls2 <- apriori (sls, parameter = list(supp = 0.002, conf = 0.6)) 

rls2

```


```{r}
# Getting the summary of the rules created
# 
summary(rls2)
```


```{r}
# Observing rules built in our model
# 
inspect(rls2[1:5])

```



# Recommendations
From the analysis the following recommendation can be made, the supermarket should stock more tea, wheat, mineral, fat, and yogurt since they are the most frequent items sold.

