---
title: "Feature Selection IP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Defining the Question?
## Specifying the Question
You are a Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax)

Perform Feature Selection.

## Defining the metrics of success
Perform data cleaning, Exploratory data analysis and Feature Selection effectively in R

## Understanding the Context
Feature selection helps selection select the important features in a dataset. This helps businesses know which features to put more importance on and how they affect their businesses. Feature selection also help reduce features needed for analysis.

## Recording the Experimental Design
 - Reading the Data
 - Checking the Data
 - Tidying the Data
 - Exploratory Data Analysis
 - Performing Feature Selection

## Data Relevance
The dataset provided was relevant for the objective we set for our project.



# Reading the Dataset
```{r}
# Loading the Dataset
sup = read.csv('Supermarket_Dataset_1 - Sales Data.csv')
```



# Checking the Dataset
```{r}
# Checking the shape of the dataset
dim(sup)
```
The dataset contains 1000 rows and 10 columns.


```{r}
# Previewing the top of the Dataset
head(sup)
```


```{r}
# Previewing the bottom of the Dataset
tail(sup)
```


```{r}
# Checking the Datatypes of the columns
str(sup)
```
The dataset contains 1 column of datatype integer, 7 columns of datatype num and 8 columns of datatype character.


```{r}
# Checking unique values of Branch
unique(sup$Branch)
```

```{r}
# Checking unique Customer type
unique(sup$Customer.type)
```

```{r}
# Checking unique values of Product line
unique(sup$Product.line)
```

```{r}
# Checking unique values of Payment
unique(sup$Payment)
```



```{r}
# Checking for Duplicates
anyDuplicated(sup)
```
There were no duplicates in our dataset.



# Tidying the Dataset
## Missing values
```{r}
# Checking for missing values
colSums(is.na(sup))
```


## Outliers
```{r}
# Checking for Outliers in Unit price column
boxplot(sup$Unit.price)
```


```{r}
# Checking for Outliers in Quantity column
boxplot(sup$Quantity)
```


```{r}
# Checking for Outliers in Tax column
boxplot(sup$Tax)
```


```{r}
# Checking for Outliers in Cogs column
boxplot(sup$cogs)
```


```{r}
# Checking for Outliers in Gross margin percentage column
boxplot(sup$gross.margin.percentage)
```


```{r}
# Checking for Outliers in Gross income column
boxplot(sup$gross.income)
```


```{r}
# Checking for Outliers in Rating column
boxplot(sup$Rating)
```


```{r}
# Checking for Outliers in Total column
boxplot(sup$Total)
```


```{r}
# Checking the list of Outliers in Tax column
boxplot.stats(sup$Tax)$out
```


```{r}
# Checking the list of Outliers in Cogs column
boxplot.stats(sup$cogs)$out
```


```{r}
# Checking the list of Outliers in Total column
boxplot.stats(sup$Total)$out
```
Most columns have 9 Outliers
The Outliers in the various columns were not removed because there were few and could not have any effect on the analysis.



# Exploratory Data Analysis
## Univariate Analysis
### Counts
```{r}
# Checking the counts of Branch 
library(plyr)
count(sup$Branch)
```


```{r}
# Checking the counts of Customer type
count(sup$Customer.type)
```


```{r}
# Checking the counts of Gender
count(sup$Gender)
```


```{r}
# Checking the counts of Product line
count(sup$Product.line)
```


```{r}
# Checking the counts of Payment
count(sup$Payment)
```


### Mean
```{r}
mean(sup$Unit.price)
```


```{r}
mean(sup$Quantity)
```


```{r}
mean(sup$Tax)
```


```{r}
mean(sup$cogs)
```


```{r}
mean(sup$gross.income)
```


```{r}
mean(sup$Rating)
```


```{r}
mean(sup$Total)
```


### Median
```{r}
median(sup$Unit.price)
```


```{r}
median(sup$Quantity)
```


```{r}
median(sup$Tax)
```


```{r}
median(sup$cogs)
```


```{r}
median(sup$gross.income)
```


```{r}
median(sup$Rating)
```


```{r}
median(sup$Total)
```


### Maximum
```{r}
max(sup$Unit.price)
```


```{r}
max(sup$Quantity)
```


```{r}
max(sup$Tax)
```


```{r}
max(sup$cogs)
```


```{r}
max(sup$gross.income)
```


```{r}
max(sup$Rating)
```


```{r}
max(sup$Total)
```


### Minimum
```{r}
min(sup$Unit.price)
```


```{r}
min(sup$Quantity)
```


```{r}
min(sup$Tax)
```


```{r}
min(sup$cogs)
```


```{r}
min(sup$gross.income)
```


```{r}
min(sup$Rating)
```


```{r}
min(sup$Total)
```


### Range
```{r}
range(sup$Unit.price)
```


```{r}
range(sup$Quantity)
```


```{r}
range(sup$Tax)
```


```{r}
range(sup$cogs)
```


```{r}
range(sup$gross.income)
```


```{r}
range(sup$Rating)
```


```{r}
range(sup$Total)
```


### Quantile
```{r}
quantile(sup$Unit.price)
```


```{r}
quantile(sup$Quantity)
```


```{r}
quantile(sup$Tax)
```


```{r}
quantile(sup$cogs)
```


```{r}
quantile(sup$gross.income)
```


```{r}
quantile(sup$Rating)
```


```{r}
quantile(sup$Total)
```


### Interquantile Range
```{r}
IQR(sup$Unit.price)
```


```{r}
IQR(sup$Quantity)
```


```{r}
IQR(sup$Tax)
```


```{r}
IQR(sup$cogs)
```


```{r}
IQR(sup$gross.income)
```


```{r}
IQR(sup$Rating)
```


```{r}
IQR(sup$Total)
```


### Variance
```{r}
var(sup$Unit.price)
```


```{r}
var(sup$Quantity)
```


```{r}
var(sup$Tax)
```


```{r}
var(sup$cogs)
```


```{r}
var(sup$gross.income)
```


```{r}
var(sup$Rating)
```


```{r}
var(sup$Total)
```


### Standard Deviation
```{r}
sd(sup$Unit.price)
```


```{r}
sd(sup$Quantity)
```


```{r}
sd(sup$Tax)
```


```{r}
sd(sup$cogs)
```


```{r}
sd(sup$gross.income)
```


```{r}
sd(sup$Rating)
```


```{r}
sd(sup$Total)
```


### Skewness
```{r}
library(e1071)
skewness(sup$Unit.price)
```


```{r}
skewness(sup$Quantity)
```


```{r}
skewness(sup$Tax)
```


```{r}
skewness(sup$cogs)
```


```{r}
skewness(sup$gross.income)
```


```{r}
skewness(sup$Rating)
```


```{r}
skewness(sup$Total)
```


### Kurtosis
```{r}
kurtosis(sup$Unit.price)
```


```{r}
kurtosis(sup$Quantity)
```


```{r}
kurtosis(sup$Tax)
```


```{r}
kurtosis(sup$cogs)
```


```{r}
kurtosis(sup$gross.income)
```


```{r}
kurtosis(sup$Rating)
```


```{r}
kurtosis(sup$Total)
```


### Plots
```{r}
# 
barplot(table(sup$Branch),
main = "Bargraph of Branches",
xlab = "Branch",
ylab = "Freq",
col = cm.colors(4),
horiz = FALSE)
```


```{r}
# 
barplot(table(sup$Customer.type),
main = "Bargraph of Customer type",
xlab = "Customer type",
ylab = "Freq",
col = cm.colors(2),
horiz = FALSE)
```


```{r}
# 
barplot(table(sup$Gender),
main = "Bargraph of Gender",
xlab = "Gender",
ylab = "Freq",
col = topo.colors(2),
horiz = FALSE)
```



```{r}
# 
barplot(table(sup$Product.line),
main = "Bargraph of Product line",
xlab = "Product line",
ylab = "Freq",
col = heat.colors(6),
horiz = FALSE)
```


```{r}
# 
barplot(table(sup$Payment),
main = "Bargraph of Payment",
xlab = "Payment",
ylab = "Freq",
col = heat.colors(3),
horiz = FALSE)
```



### Histograms
```{r}
hist(sup$Unit.price)
```


```{r}
hist(sup$Quantity)
```


```{r}
hist(sup$Tax)
```


```{r}
hist(sup$cogs)
```


```{r}
hist(sup$gross.income)
```


```{r}
hist(sup$Rating)
```


```{r}
hist(sup$Total)
```


## Bivariate Analysis
### Plots
```{r}
library(ggplot2)
ggplot(data=sup, aes(x=Quantity, y=Unit.price)) +
geom_bar(stat="identity")
```


```{r}
library(ggplot2)
ggplot(data=sup, aes(x=Quantity, y=Rating)) +
geom_bar(stat="identity")
```


```{r}
# Creating a Dataframe with numeric variables only
sup1 = sup[, c(6,7,8,12,14:16)]

head(sup1)
```


### Covariance
```{r}
# Finding the covariance of the feature variables
cov(sup1)
```


### Correlation
```{r}
# Finding the correlation coefficients of feature variables 
cor(sup1)
```


### Correlation Matrix
```{r}
# Creating a correlation matrix
library('corrplot')

corrplot(cor(sup1), order = "hclust")
```


### Scatter Plots
```{r}
# Plotting a scatterplot between Cogs and Quantity
plot(sup$cogs, sup$Quantity, xlab="Cogs", ylab="Quantity")
```


```{r}
# Plotting a scatterplot between Tax and Unit price
plot(sup$Tax, sup$Unit.price, xlab="Tax", ylab="Unit price")
```


```{r}
# Plotting a scatterplot between Rating and Gross income
plot(sup$Rating, sup$gross.income, xlab="Rating", ylab="Gross income")
```


```{r}
# Plotting a scatterplot between Rating and Total
plot(sup$Rating, sup$Total, xlab="Rating", ylab="Total")
```


### Observations
 - The branch with the most customers was Branch A.
 - Most customers were members and others normal customers.
 - Most customers were females compared to males.
 - Fashion accessories were the most bought products.
 - Most customers payed using Ewallets.
 - Most customers bought 1 - 2 items per sale.
 - Most items had taxes ranging from 0 - 15.
 - Most items had cogs between 0 - 200.
 - Most customers rated 4.
 - Tax, Cogs, and Total are highly correlated according to the correlation matrix.



# Implementing the Solution
## Feature Selection
### Filter Method
```{r}
# Creating a numeric only dataset
sp = sup[, c(6,7,8,12,14:16)]
head(sp)

```


```{r}
# Loading the required libraries
#
library(caret)
library(corrplot)

# Getting the correlations of the variables.
#
correlationMatrix <- cor(sp)
```


```{r}
# Find attributes that are highly correlated
#
hc <- findCorrelation(correlationMatrix, cutoff=0.75)

# Showing the highly correlated attributes
#
names(sp[,hc])
```


```{r}
# Removing Redundant Features
# 
sp2<-sp[-hc]

# Performing our graphical comparison
# 
par(mfrow = c(1, 2))
corrplot(correlationMatrix, order = "hclust")
corrplot(cor(sp2), order = "hclust")
```


### Embedded Method
```{r}
# Loading the required library
#
library(wskm)

# Creating a model
#
set.seed(2)
model <- ewkm(sp, 3, lambda=2, maxiter=1000)
```


```{r}
# Loading the required library
#
library("cluster")

# Cluster Plot against 1st and 2nd principal components
#
clusplot(sp, model$cluster, color=TRUE, shade=TRUE, labels=3, lines=1, main='Cluster Analysis for Sales')

```


```{r}
# Getting the weights in percentages
round(model$weights*100,2)
```


### Wrapper Method
```{r}
# Loading the required libraries
#
library(clustvarsel)
library(mclust)

# Performing sequential forward headlong search
#
out = clustvarsel(sp, G = 1:5, search = c("headlong"))
out

```


```{r}
# Building the clustering model
#
sp_sub = sp[,out$subset]
mod = Mclust(sp_sub, G = 1:5)
summary(mod)

```


```{r}
# Plotting the model
#
plot(mod,c("classification"))

```


## Observations
 - From Filter method we can see that tax, cogs, and total are highly correlated and therefore should be removed.
 - From the Wrapper method tax and gross income are selected to be added and the other variables rejected.
 

# Recommendations
From the analysis the following recommendations can be made the filter method is the best to use for feature selection since it removes features which are highly correlated. Tax and gross income are the best feature to consider in our dataset when doing analysis. 

